The impact of cosmic variance on simulating weak lensing surveys

### 2014-09-26 Writeup/paper_v3.tex
Finding ellip methods swapped

%\newcommand{\rachel}[1]{{\textcolor{red}{#1}}}
\newcommand{\rachel}[1]{}
%\newcommand{\arun}[1]{{\textcolor{blue}{#1}}}
\newcommand{\arun}[1]{}
\def\a{\alpha}
\def\reff@jnl#1{{\rm#1\/}}

@@

% \rachel{Overall comment: the figures need a {\em lot} of work.  First,  they should be in the same order as they are referenced in the  text.  Second, they need to be sensibly referenced in the text; you  can't say ``as we see\dots'' without having a clear statement that  we should be looking at Fig.~4 (or whatever).  In all  cases, the font size is way too small, so you should enlarge it.  The rule of thumb for MNRAS and other journals is that the font size  in the figures should appear the same size as the font in the text  of the paper.  (This is true for axis labels, tick labels, and text  in the legend.)  You also need to write good captions, because some  people will not really read the paper, just the captions.  MNRAS
%         does not use the subfigure environment, so you should remove it.  Just have the figures on top of or next to each other, and then in  the caption refer to ``Top panel'' and ``Bottom panel'' or whatever.} 
%         \arun{Results have too many figures and they overlap with the bibliography. I think we should put some of the figures explicitly after the references?}
% \rachel{No, just move the figures a bit earlier in the text.  It's okay for them to show up slightly before you actually discuss them.  The references really should be last.}
\section{Introduction}
\label{S:intro}

@@

\arun{Talk more about environments - Nature and nurture?}
\rachel{What do you have in mind here?}

@@

If galaxies have elliptical isophotes, its shape and size could be defined by the axis ratio and the area enclosed by a boundary isophote. However, in real galaxy images,
the boundary may not be well defined and the shape may not be well approximated by an ellipse. More importantly, the effect of smearing due to the point-spread function (PSF) is larger. 
% Given the surface brightness (intensity) $I(\vec{\theta})$ of a galaxy image at every angular position $\vec{\theta}$,
% we can define the tensor of second brightness moments as
% \begin{equation}
%  Q_{ij} = \frac{\int\rmd^2\theta q_I (\theta_i - \overline{\theta_i})(\theta_j - \overline{\theta_j})}{\int\rmd^2\theta q_I},
% \end{equation}
% where $q_I$ is a suitable chosen kernel and $\overline{\vec{\theta}}$ is the average angular coordinate, weighted by the same kernel function and $i,j\in\lbrace 1,2 \rbrace$ are the components of $\vec{\theta}$.
% It is common to find in literature two kinds of \emph{complex }ellipticities that quantify the shape of the galaxies - $\chi$ and $\epsilon$ defined as
% \begin{align}
%  \chi &= \frac{Q_{11}-Q_{12}+2iQ_{12}}{Q_{11}+Q_{22}} \\ \epsilon &= \frac{Q_{11}-Q_{22}+2iQ_{12}}{Q_{11}+Q_{22}+2(Q_{11}Q_{22}-Q_{12})^{1/2}}
% \end{align}
% If the image has elliptical isophotes with axis ratio $q$, then 
% \begin{equation}
%  |\chi| = \frac{1-q^2}{1+q^2}, \;\;\; |\epsilon| = \frac{1-q}{1+q}. 
% \end{equation}
% 
% 
% Conversely, one could compute $\chi$ or $\epsilon$ for a galaxy, and obtain an effective, azimuthally averaged {\bf(?)} axis-ratio $q$. 
% \rachel{No, that isn't possible because of the PSF.  One would have to
%   use a PSF correction scheme.  Recommend removing much of the above discussion.}
% \arun{Should I remove these even if we plan to use the ellipticities from the second moments?}

The PSF correction scheme used on the observed images is that of `re-Gaussianization' method described in \S 2.4 of~\cite{HS03} (hereafter HS03).
This method models the true PSF $g({\bf x})$ as a Gaussian $G({\bf x})$ and the residual $\epsilon({\bf x}) = g({\bf x}) - G({\bf x})$ is assumed to be small. Thus, the Gaussian-convolved
intrinsic image of $f$, is $I' = G\otimes f = I - \epsilon \otimes f$, where $I$ is the observed image. The crucial idea here is that, when $\epsilon$ is small, we get a reasonably accurate
estimate of $I'$ even if we use an approximate form for $f$. The form assumed for $f$ is that of a Gaussian with covariance $M_f^{(0)} = M_I - M_g$, where $M_I$ and $M_g$ are the adaptive
covariances of the measured object and PSF respectively, described in \S 2.1 of HS03, which is in turn based on~\cite{BJ02}. Once we obtain the covariance matrices of the instrinsic image $f$, one can compute the ellipticities 
of the galaxies, which we will refer to as `ellipticities based on moments'.

Another method of estimate the ellipticities is to fit parametric models to each image. Two examples of the models are
\begin{enumerate}
 \item a \sersic profile given by the expression 
       \begin{equation} 
    S = \Sigma_{1/2}\exp{\left( -k(R/R_{\text{eff}})^{1/n} -1 \right)},
       \end{equation}
       \item two \sersic component fits: de Vaucoulers bulge $(n=4)$ + exponential disc profile $(n=1)$,
\end{enumerate}
where \begin{align*} R^2 = & ((x-x_0)\cos\Phi+(y-y_0)\sin\Phi)^2  \\ & + ((y-y_0)\cos\Phi-(x-x_0)\sin\Phi)^2/q^2, \end{align*}
$R_{\text{eff}}$ is the half-light radius of the profile, $\Sigma_{1/2}$ is the surface brightness at $R=R_{\text{eff}}$, $(x_0,y_0)$ is the centroid of the image,
$\Phi$ is the profile rotation angle, $n$ is the \sersic index, $k$ is a $n$-dependent normalization factor and $q$ is the axis ratio of the elliptical isophotes.
Thus, the \sersic profile has 7 free parameters. The bulge+disk model has 10 free parameters since the \sersic indices are fixed as 1 and 4
and both the profiles are required to have the same centroid $(x_0,y_0)$. Best-fit parameters were found, by minimizing the weighted
sum of the difference between the image and the PSF-convolved mode
using Levenberg-Marquardt minimization, \texttt{mpfit2fun} in IDL~\citep{mpfit2fun}. More details about the fit can be found
in~\cite{Claire_Fits}.

@@

@@

\subsection{Axis-ratios} 
\begin{figure*}
 \centering
 \includegraphics[width=2.3\columnwidth]{axis_ratio_all}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all.eps}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all_Bbandevolution.eps}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all_masscut.eps}
 \caption{The distributions of axis ratios of galaxies in \emph{all} overdense (OD) and \emph{all} underdense (UD) regions in the case of luminosity-selected sample (left), luminosity-selected with B-band evolution taken into account (center) and stellar-mass-selected sample (right). The upper panels show the histogram
 and the botom panels show the empirical cumulative distribution function (ECDF). $p$-values are computed using these CDFs using the Kolmogorov-Smirnov and Anderson-Darling tests and are given in Table. \ref{table:pvalues_all}. The difference between the CDFs turn out to be statistically significant.}
 \label{fig:axisratio_all}
\end{figure*}

@@

We can understand the influence of environment on the ellipticities of the galaxies simply by comparing the distribution of the axis ratios for the overdense and underdense samples. 
%The distribution of the axis ratios that are obtained by the above manner are compared between redshift bins. 
We use two statistical tests namely the Kolmogorov-Smirnov test and Anderson-Darling test to compare distributions.

@@

\begin{figure}
 %\centering
 a) \includegraphics[width=1.0\columnwidth]{rms_ellip1_noevolution.eps} \
 b) \includegraphics[width=1.0\columnwidth]{rms_ellip2_noevolution.eps} \\
 \caption{Left: RMS ellipticities with ellipticity defined as $\frac{1-q}{1+q}$. \; 
          Right: RMS ellipticities with ellipticity defined as $\frac{1-q^2}{1+q^2}$.\\ The horizontal errorbars simply correspond to the binwidth while the vertical ones
          are $1\sigma$ errorbars obtained by bootstrapping.}
 \label{fig:rms_ellip}
\end{figure}

@@

%The shape of a population of galaxies can be characterized by a single number called `RMS ellipticity'. If $q=b/a$ is the axis-ratio, then the ellipticity maybe defined as $\frac{1-q}{1+q}$
%or as $\frac{1-q^2}{1+q^2}$. The root mean squared (RMS) of ellipticities of galaxies in each redshift bin are shown in Figure~\ref{fig:rms_ellip}.
It is useful to consider the root mean square ellipticity, a statistic that can characterize the shape of the population/sample of galaxies.
For the luminosity-selected sample without any evolution, RMS ellipticities of galaxies in each redshift bin are shown in Figure~\ref{fig:rms_ellip}. As one can see, the underdense regions
have significantly higher values for RMS ellipticities when compared to the overdense regions. Note in particular that we've been able to capture the narrow overdense bin $0.75\le z < 0.80$.  
There is (almost) no redshift dependence in the figure.

@@

When the B-band luminosity evolution is taken into account in selecting the sample, a systematic increase in the ellipticity at lower redshifts can be observed. We plot these alongside with stellar-mass selected samples, where a similar trend is found, in Fig.~\ref{fig:rms_ellip_Bband_mass}. 
Such a trend is not observed in axis ratios based on second moments. 
For other morphological parameters, we do not the distributions themselves directly. For \sersic profile fits, it gets tricky since we truncate the \sersic index at 6. 
For \btt ratios, we force the value to be greater than 0.05 \arun{What exactly do we do here?}. We will understand the dependence of these quantities on environment by computing the median values in different redshift bins.
We choose median over mean since it is more robust to the truncation effects.
Figure(s) \ref{fig:median_sersicn} show the median of the \sersic index with and without taking into account of the luminosity evolution. Median values obtained using stellar mass selected samples are plotted alongside for reference.
We observe that the overdense regions tend to have higher \sersic index than the adjacent underdense ones.
Figure(s) \ref{fig:median_dvcbtt} show the variation of median \btt ratio with redshift. We see that the bulge component is more in overdense regions than in the underdense regions.
These two observations are consistent with each other since higher \sersic index implies higher bulge component which are typical in young galaxies found in overdense regions.
%We chose median over mean since \sersic n greater than 6 are truncated to 6. This affects the mean statistic more than the median.
\begin{figure}
 %\centering
 \includegraphics[width=1.1\columnwidth]{median_sersicn}
 \includegraphics[width=1.1\columnwidth]{median_sersicn2}
 \caption{Median values of the \sersic indices for volume-limited samples \s$1$ and \s$3$ are plotted (filled centers and thin errorbars) in left and right panels respectively for each redshift bin.
          Median values for the \s$2$ sample are plotted in both the panels (open centers and thick errorbars) in both the panels.
          The horizontal errorbars simply correspond to the binwidth while the vertical ones are $1\sigma$ errorbars obtained by bootstrapping.}
 \label{fig:median_sersicn}
\end{figure}

\begin{figure}
 %\centering
 \includegraphics[width=1.1\columnwidth]{median_dvcbtt1}
 \includegraphics[width=1.1\columnwidth]{median_dvcbtt2}
  \caption{Median values of the \btt ratios for volume-limited samples \s$1$ and \s$3$ are plotted (filled centers and thin errorbars) in left and right panels respectively for each redshift bin.
          Median values for the \s$2$ sample are plotted in both the panels (open centers and thick errorbars) in both the panels.
          The horizontal errorbars simply correspond to the binwidth while the vertical ones are $1\sigma$ errorbars obtained by bootstrapping.}
 \label{fig:median_dvcbtt}
\end{figure}

>>>>

\newcommand{\rachel}[1]{{\textcolor{red}{#1}}}
%\newcommand{\rachel}[1]{}
\newcommand{\arun}[1]{{\textcolor{blue}{#1}}}
%\newcommand{\arun}[1]{}
\def\a{\alpha}
\def\reff@jnl#1{{\rm#1\/}}

@@

\rachel{Overall comment: the figures need a {\em lot} of work.  First,  they should be in the same order as they are referenced in the  text.  Second, they need to be sensibly referenced in the text; you  can't say ``as we see\dots'' without having a clear statement that  we should be looking at Fig.~4 (or whatever).  In all  cases, the font size is way too small, so you should enlarge it.  The rule of thumb for MNRAS and other journals is that the font size  in the figures should appear the same size as the font in the text  of the paper.  (This is true for axis labels, tick labels, and text  in the legend.)  You also need to write good captions, because some  people will not really read the paper, just the captions.  MNRAS
         does not use the subfigure environment, so you should remove it.  Just have the figures on top of or next to each other, and then in  the caption refer to ``Top panel'' and ``Bottom panel'' or whatever.} 
         \arun{Results have too many figures and they overlap with the bibliography. I think we should put some of the figures explicitly after the references?}
 \rachel{No, just move the figures a bit earlier in the text.  It's okay for them to show up slightly before you actually discuss them.  The references really should be last.}
\section{Introduction}
\label{S:intro}

@@

%\arun{Talk more about environments - Nature and nurture?}
%\rachel{What do you have in mind here?}

@@

If galaxies have elliptical isophotes, its shape and size could be defined by the axis ratio and the area enclosed by a boundary isophote. However, in real galaxy images,
the boundary may not be well defined and the shape may not be well approximated by an ellipse. More importantly, the effect of smearing due to the point-spread function (PSF) is larger. 
Thus, we are in need of more sophisticated methods to measure the ellipticities.
% Given the surface brightness (intensity) $I(\vec{\theta})$ of a galaxy image at every angular position $\vec{\theta}$,
% we can define the tensor of second brightness moments as
% \begin{equation}
%  Q_{ij} = \frac{\int\rmd^2\theta q_I (\theta_i - \overline{\theta_i})(\theta_j - \overline{\theta_j})}{\int\rmd^2\theta q_I},
% \end{equation}
% where $q_I$ is a suitable chosen kernel and $\overline{\vec{\theta}}$ is the average angular coordinate, weighted by the same kernel function and $i,j\in\lbrace 1,2 \rbrace$ are the components of $\vec{\theta}$.
% It is common to find in literature two kinds of \emph{complex }ellipticities that quantify the shape of the galaxies - $\chi$ and $\epsilon$ defined as
% \begin{align}
%  \chi &= \frac{Q_{11}-Q_{12}+2iQ_{12}}{Q_{11}+Q_{22}} \\ \epsilon &= \frac{Q_{11}-Q_{22}+2iQ_{12}}{Q_{11}+Q_{22}+2(Q_{11}Q_{22}-Q_{12})^{1/2}}
% \end{align}
% If the image has elliptical isophotes with axis ratio $q$, then 
% \begin{equation}
%  |\chi| = \frac{1-q^2}{1+q^2}, \;\;\; |\epsilon| = \frac{1-q}{1+q}. 
% \end{equation}
% 
% 
% Conversely, one could compute $\chi$ or $\epsilon$ for a galaxy, and obtain an effective, azimuthally averaged {\bf(?)} axis-ratio $q$. 
% \rachel{No, that isn't possible because of the PSF.  One would have to
%   use a PSF correction scheme.  Recommend removing much of the above discussion.}
% \arun{Should I remove these even if we plan to use the ellipticities from the second moments?}

One method to estimate the ellipticities is to estimate the axis-ratios by fitting parametric models to each image. The models that we fit to the images are
\begin{enumerate}
 \item a \sersic profile given by the expression 
       \begin{equation} 
    S = \Sigma_{1/2}\exp{\left( -k(R/R_{\text{eff}})^{1/n} -1 \right)},
       \end{equation}
       \item two \sersic component fits: de Vaucoulers bulge $(n=4)$ + exponential disc profile $(n=1)$,
\end{enumerate}
where \begin{align*} R^2 = & ((x-x_0)\cos\Phi+(y-y_0)\sin\Phi)^2  \\ & + ((y-y_0)\cos\Phi-(x-x_0)\sin\Phi)^2/q^2, \end{align*}
$R_{\text{eff}}$ is the half-light radius of the profile, $\Sigma_{1/2}$ is the surface brightness at $R=R_{\text{eff}}$, $(x_0,y_0)$ is the centroid of the image,
$\Phi$ is the profile rotation angle, $n$ is the \sersic index, $k$ is a $n$-dependent normalization factor and $q$ is the axis ratio of the elliptical isophotes.
Thus, the \sersic profile has 7 free parameters. The bulge+disk model has 10 free parameters since the \sersic indices are fixed as 1 and 4
and both the profiles are required to have the same centroid $(x_0,y_0)$. Best-fit parameters were found, by minimizing the weighted
sum of the difference between the image and the PSF-convolved model
using Levenberg-Marquardt minimization, \texttt{mpfit2fun} in IDL~\citep{mpfit2fun}. More details about the fit can be found
in~\cite{Claire_Fits}.

@@

An alternative method involves correcting the observed image for the PSF and computing the covariances from which the ellipticity of the galaxy can be obtained.
The PSF correction scheme used on the observed images is that of `re-Gaussianization' method described in \S 2.4 of~\cite{HS03} (hereafter HS03).
This method models the true PSF $g({\bf x})$ as a Gaussian $G({\bf x})$ and the residual $\epsilon({\bf x}) = g({\bf x}) - G({\bf x})$ is assumed to be small. Thus, the Gaussian-convolved
intrinsic image of $f$, is $I' = G\otimes f = I - \epsilon \otimes f$, where $I$ is the observed image. The crucial idea here is that, when $\epsilon$ is small, we get a reasonably accurate
estimate of $I'$ even if we use an approximate form for $f$. The form assumed for $f$ is that of a Gaussian with covariance $M_f^{(0)} = M_I - M_g$, where $M_I$ and $M_g$ are the adaptive
covariances of the measured object and PSF respectively, described in \S 2.1 of HS03, which is in turn based on~\cite{BJ02}. Once we obtain the covariance matrices of the instrinsic image $f$, one can compute the ellipticities 
of the galaxies, which we will refer to as `ellipticities based on moments'.

@@

\subsection{Axis-ratios} 
We can understand the influence of environment on the ellipticities of the galaxies simply by comparing the distribution of the axis ratios for the overdense and underdense samples. 
Then, we proceed to consider the root mean square ellipticity, a statistic that can characterize the shape of the population/sample of galaxies. 
\subsubsection{Comparing distributions}
\begin{figure*}
 \centering
 \includegraphics[width=2.3\columnwidth]{axis_ratio_all}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all.eps}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all_Bbandevolution.eps}
 %\includegraphics[width=0.66\columnwidth]{axis_ratio_all_masscut.eps}
 \caption{The distributions of axis ratios of galaxies in \emph{all} overdense (OD) and \emph{all} underdense (UD) regions in the case of luminosity-selected sample (left), luminosity-selected with B-band evolution taken into account (center) and stellar-mass-selected sample (right). The upper panels show the histogram
 and the botom panels show the empirical cumulative distribution function (ECDF). $p$-values are computed using these CDFs using the Kolmogorov-Smirnov and Anderson-Darling tests and are given in Table. \ref{table:pvalues_all}. The difference between the CDFs turn out to be statistically significant.}
 \label{fig:axisratio_all}
\end{figure*}

@@

%The distribution of the axis ratios that are obtained by the above manner are compared between redshift bins. 
We use two statistical tests namely the Kolmogorov-Smirnov test and Anderson-Darling test to compare distributions.

@@

\subsubsection{RMS ellipticites}
\begin{figure}
 %\centering
 a) \includegraphics[width=1.0\columnwidth]{rms_ellip1_noevolution.eps} \
 b) \includegraphics[width=1.0\columnwidth]{rms_ellip2_noevolution.eps} \\
 \caption{Left: RMS ellipticities with ellipticity defined as $\frac{1-q}{1+q}$. \; 
          Right: RMS ellipticities with ellipticity defined as $\frac{1-q^2}{1+q^2}$.\\ The horizontal errorbars simply correspond to the binwidth while the vertical ones
          are $1\sigma$ errorbars obtained by bootstrapping.}
 \label{fig:rms_ellip}
\end{figure}

@@

%The shape of a population of galaxies can be characterized by a single number called `RMS ellipticity'. If $q=b/a$ is the axis-ratio, then the ellipticity maybe defined as $\frac{1-q}{1+q}$
%or as $\frac{1-q^2}{1+q^2}$. The root mean squared (RMS) of ellipticities of galaxies in each redshift bin are shown in Figure~\ref{fig:rms_ellip}.
For the luminosity-selected sample without any evolution, RMS ellipticities of galaxies in each redshift bin are shown in Figure~\ref{fig:rms_ellip}. As one can see, the underdense regions
have significantly higher values for RMS ellipticities when compared to the overdense regions. Note in particular that we've been able to capture the narrow overdense bin $0.75\le z < 0.80$.  
There is (almost) no redshift dependence in the figure.

@@

When the B-band luminosity evolution is taken into account in selecting the sample, a systematic increase in the ellipticity at lower redshifts can be observed. We plot these alongside with stellar-mass selected samples, where a similar trend is found, in Fig.~\ref{fig:rms_ellip_Bband_mass}. 
Such a trend is not observed in axis ratios based on second moments.

\subsection{\sersicn}
For other morphological parameters, we do not compare the distributions themselves directly. For \sersic profile fits, it gets tricky since we truncate the \sersic index at 6. 
For \btt ratios, we force the value to be greater than 0.05 \arun{What exactly do we do here?}. We will understand the dependence of these quantities on environment by computing the median values in different redshift bins.
We choose median over mean since it is more robust to the truncation effects.
Figure(s) \ref{fig:median_sersicn} show the median of the \sersic index with and without taking into account of the luminosity evolution. Median values obtained using stellar mass selected samples are plotted alongside for reference.
We observe that the overdense regions tend to have higher \sersic index than the adjacent underdense ones.
Figure(s) \ref{fig:median_dvcbtt} show the variation of median \btt ratio with redshift. We see that the bulge component is more in overdense regions than in the underdense regions.
These two observations are consistent with each other since higher \sersic index implies higher bulge component which are typical in young galaxies found in overdense regions.
%We chose median over mean since \sersic n greater than 6 are truncated to 6. This affects the mean statistic more than the median.
\begin{figure}
 %\centering
 \includegraphics[width=1.0\columnwidth]{median_sersicn}
 \includegraphics[width=1.0\columnwidth]{median_sersicn2}
 \caption{Median values of the \sersic indices for volume-limited samples \s$1$ and \s$3$ are plotted (filled centers and thin errorbars) in left and right panels respectively for each redshift bin.
          Median values for the \s$2$ sample are plotted in both the panels (open centers and thick errorbars) in both the panels.
          The horizontal errorbars simply correspond to the binwidth while the vertical ones are $1\sigma$ errorbars obtained by bootstrapping.}
 \label{fig:median_sersicn}
\end{figure}

\begin{figure}
 %\centering
 \includegraphics[width=1.0\columnwidth]{median_dvcbtt1}
 \includegraphics[width=1.0\columnwidth]{median_dvcbtt2}
  \caption{Median values of the \btt ratios for volume-limited samples \s$1$ and \s$3$ are plotted (filled centers and thin errorbars) in left and right panels respectively for each redshift bin.
          Median values for the \s$2$ sample are plotted in both the panels (open centers and thick errorbars) in both the panels.
          The horizontal errorbars simply correspond to the binwidth while the vertical ones are $1\sigma$ errorbars obtained by bootstrapping.}
 \label{fig:median_dvcbtt}
\end{figure}
